---
title: "Homework 2"
author: "Nicole Magallanes"
subtitle: PSTAT 131/231
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    toc_float: yes
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE,
                      warning = FALSE)
```

## Linear Regression
```{r}
library(tidymodels)
library(tidymodels)

abalone <- read.csv("/Users/nicolemagallanes/Desktop/hw2-nicolemagallanes/abalone.csv")
#view(abalone)
```


### Question 1

Your goal is to predict abalone age, which is calculated as the number of rings plus 1.5. Notice there currently is no `age` variable in the data set. Add `age` to the data set.

Assess and describe the distribution of `age`.

```{r}
age <- abalone$rings + 1.5
abalone$age <- age
# view(abalone)

abalone %>%
  ggplot(aes(x=age)) +
  geom_histogram(bins=40)
```

The distribution of Age seems to be somewhat positively skewed.We can see the data peaks at around age of 11-12 and as the distribution is right skewed we can see that there aren't a lot of abalone that are past the age of 20. 

### Question 2

Split the abalone data into a training set and a testing set. Use stratified sampling. You should decide on appropriate percentages for splitting the data.

*Remember that you'll need to set a seed at the beginning of the document to reproduce your results.*

```{r}
set.seed(4160)

abalone_split <- initial_split(abalone, prop = 0.80,
                                strata = age)
abalone_train <- training(abalone_split)
abalone_test <- testing(abalone_split)

#abalone_train
#abalone_test
```

### Question 3

Using the **training** data, create a recipe predicting the outcome variable, `age`, with all other predictor variables. Note that you should not include `rings` to predict `age`. Explain why you shouldn't use `rings` to predict `age`.

Steps for your recipe:

1.  dummy code any categorical predictors

2.  create interactions between

    -   `type` and `shucked_weight`,
    -   `longest_shell` and `diameter`,
    -   `shucked_weight` and `shell_weight`

3.  center all predictors, and

4.  scale all predictors.

You'll need to investigate the `tidymodels` documentation to find the appropriate step functions to use.

```{r}
abalone_recipe <- recipe(age~., data= abalone_train) %>%
  step_rm(rings) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_interact(terms = ~ starts_with("type"):shucked_weight +
                  longest_shell:diameter + shucked_weight:shell_weight) %>% 
 step_scale(all_numeric_predictors()) %>%
  step_center(all_numeric_predictors())
abalone_recipe
```
We do not include rings because rings to predict age because rings is already use to see the age, it is a variable that is included withing age. 

### Question 4

Create and store a linear regression object using the `"lm"` engine.
```{r}
lm_model <- linear_reg() %>%
  set_engine("lm")
```

### Question 5

Now:

1.  set up an empty workflow,
2.  add the model you created in Question 4, and
3.  add the recipe that you created in Question 3.

```{r}
lm_wflow <- workflow() %>%
  add_model(lm_model) %>%
  add_recipe(abalone_recipe)
```


### Question 6

Use your `fit()` object to predict the age of a hypothetical female abalone with longest_shell = 0.50, diameter = 0.10, height = 0.30, whole_weight = 4, shucked_weight = 1, viscera_weight = 2, shell_weight = 1.

```{r}
lm_fit <- fit(lm_wflow, abalone_train)
#lm_fit

newd <- data.frame(type = "F", longest_shell = 0.50, diameter = 0.10, height = 0.30, whole_weight = 4, shucked_weight = 1, viscera_weight = 2, shell_weight = 1, rings = 0)

#view(newd)

predict(lm_fit, new_data = newd)
```

### Question 7

Now you want to assess your model's performance. To do this, use the `yardstick` package:

1.  Create a metric set that includes *R^2^*, RMSE (root mean squared error), and MAE (mean absolute error).
2.  Use `predict()` and `bind_cols()` to create a tibble of your model's predicted values from the **training data** along with the actual observed ages (these are needed to assess your model's performance).
3.  Finally, apply your metric set to the tibble, report the results, and interpret the *R^2^* value.

```{r}
library(yardstick)

abalone_train_res <- predict(lm_fit, new_data = abalone_train %>% select(-age))

abalone_train_res <- bind_cols(abalone_train_res, abalone_train %>% select(age))
abalone_train_res %>% 
  head()

rmse(abalone_train_res, truth = age, estimate = .pred)

abalone_metrics <- metric_set(rmse, rsq, mae)
abalone_metrics(abalone_train_res, truth = age, 
                estimate = .pred)

```
Our R-squared value tells us that 55.73% of our variance in the dependent variable can be explained by the independent variables. The higher the R-square, usually means the better our model fits the data. In this case our R-square value is really low, meaning our model only explains 55.73% of our fitted data. 
